---
title: "Paid Media ROI Reallocation (Simulated)"
format: html
---

### Question
If I have $20,000 this month, how should I reallocate across campaigns to maximize profit?

### Approach (quick version)
- Simulated dataset of 8 campaigns across 4 channels  
- ROI = (revenue - spend) / spend  
- Allocate $20k to the top quartile by ROI (proportional), then estimate impact

```{r}
library(tidyverse)

set.seed(42)
ads <- tibble(
  campaign = c("82 Sweater TikTok","61 Vest Facebook","92 Pant Instagram","32 Pant Google",
               "71 Jacket TikTok","44 Shoes Facebook","27 Hat Instagram","10 Socks Google"),
  channel  = c("TikTok","Facebook","Instagram","Google","TikTok","Facebook","Instagram","Google"),
  spend    = round(runif(8, 500, 2500), 0),
  clicks   = round(runif(8, 1500, 12000), 0)
) |>
  mutate(
    conv    = round(clicks * runif(n(), 0.01, 0.04), 0),
    aov     = round(runif(n(), 45, 120), 0),
    revenue = conv * aov,
    roi     = (revenue - spend) / spend
  ) |>
  arrange(desc(roi))

ads
```

### Baseline summary
```{r}
baseline <- ads |>
  summarise(
    campaigns   = n(),
    total_spend = sum(spend),
    total_rev   = sum(revenue),
    avg_roi     = mean(roi)
  )
baseline
```

### Reallocate $20k to top quartile by ROI
```{r}
alloc <- 20000

# top quartile of campaigns by ROI (use nrow(), not n())
top_q <- ads |> slice_head(n = ceiling(nrow(ads) / 4))

# proportional weights by ROI within top quartile
weights <- top_q$roi / sum(top_q$roi)

top_alloc <- top_q |>
  mutate(
    add_spend = round(alloc * weights, 0),
    proj_rev  = revenue + add_spend * (1 + roi)
  )

proj <- ads |>
  anti_join(top_q, by = "campaign") |>
  mutate(add_spend = 0, proj_rev = revenue) |>
  bind_rows(top_alloc)

projected <- proj |>
  summarise(
    new_spend   = sum(spend + add_spend),
    new_rev     = sum(proj_rev),
    new_avg_roi = (new_rev - new_spend) / new_spend
  )
projected
```

### Visual: ROI by campaign
```{r}
ads |>
  ggplot(aes(x = reorder(campaign, roi), y = roi, fill = channel)) +
  geom_col() +
  coord_flip() +
  labs(x = "Campaign", y = "ROI", title = "Baseline ROI by Campaign")
```

### Findings
- Baseline average ROI: `r round(baseline$avg_roi, 2)`  
- Projected average ROI after reallocating $20k to the top quartile: `r round(projected$new_avg_roi, 2)`

### Decision
Shift $20k into the top quartile campaigns. Recheck in 2 weeks; add caps if ROI decays.

### What I’d test next
Ad-set holdouts, creative rotation cadence, and channel-level saturation checks.